---
output:
  pdf_document: default
  html_document: default
---

Load Dependencies
```{r warning=FALSE, echo = TRUE, results = 'hide',message=FALSE}
# Load the required packages
library(dplyr)
library(data.table)
library(magrittr)
library(lubridate)
library(scales)  
library(ggplot2)
```

### Are Central Perk’s Customers Loyal?

*Execution and Results* <br>

An important piece of understanding the loyalty of Central Perk's customers is to explore the frequency of the customer visits. We look at the frequency of customer visits taking into account the time when they became our customer to know our repeat customers.

```{r warning=FALSE}
#Loading the data and creating a single dataset from the three years data
shop<-read.csv("Central Perk Item Sales Summary 2016.csv",stringsAsFactors = F)
shop1<-read.csv("Central Perk Item Sales Summary 2017.csv",stringsAsFactors = F)
shop2<-read.csv("Central Perk Item Sales Summary 2018.csv",stringsAsFactors = F)
total <- rbind(shop,shop1,shop2)
nrow(total)
```

Removing the customers with NA value as there is no way of knowing whether they visited us again

```{r warning=FALSE}
total1<-filter(total,total$Customer.ID!='NA')
nrow(total1)
```

Getting the number of vists per customer to know our loyal customers

```{r warning=FALSE}
#Number of items per customer visits
total2<-total1 %>%
  group_by(Customer.ID,Date,Time) %>%
  summarize(number_items = n()) 

#Number of vists of a customer in a day
shop2<-total2 %>%
  group_by(Customer.ID,Date) %>%
  summarize(number_visits_day = n()) 

#Getting month and year from date as we want to know the average customer's vist frequency per month
shop2$date1<-mdy(shop2$Date)
shop2$mon<-month(shop2$date1)
shop2$yr<-year(shop2$date1)

#Total number of vists of a customer in a given month-year
shop3<-shop2 %>%
  group_by(Customer.ID,mon,yr) %>%
  summarize(number_vists = n()) 

#find the starting month for a customer
shop3$day<-1
shop3$date_new <- as.Date(with(shop3, paste(yr, mon, day,sep="-")), "%Y-%m-%d")
shop4<-shop3 %>%
  group_by(Customer.ID) %>%
  summarize(start_mnth = min(date_new))

#Fixing the ending month as August,2018
shop4$end_mnth<-max(shop3$date_new)

#Finding the total duration of our customer engagement in terms of number of months
shop4$diff<-interval(ymd(shop4$start_mnth),ymd(shop4$end_mnth)) %/% months(1)

#total number of visits by a customer from his starting month
shop5<-shop3 %>%
  group_by(Customer.ID) %>%
  summarize(total_vists=sum(number_vists)) 

#Calculating the monthly average of the number of visits by a customer to establish their loyalty
shop6<-inner_join(shop5,shop4,by="Customer.ID")
shop6$Visits_per_month<-shop6$total_vists/shop6$diff
shop6$Visits_per_month1<-ifelse(shop6$start_mnth==shop6$end_mnth,shop6$total_vists,
                                 shop6$Visits_per_month)
```

Filtering the customers based on the different frequency of their visits and counting the total number of customers who are above different frequency thresholds.

```{r warning=FALSE}
#Finding total number of unique customers
total_cust<-as.data.frame(unique(total1$Customer.ID))
total_customers<-nrow(total_cust)

# Frequency of average monthly visits greator or equal to 1
shop7<-filter(shop6,Visits_per_month1>=1)
more_than_or_equal_to_one=nrow(shop7)
p1=(more_than_or_equal_to_one/total_customers)*100
p1
```

```{r warning=FALSE}
# Frequency of average monthly visits greator or equal to 2
shop7<-filter(shop6,Visits_per_month1>=2)
more_than_or_equal_to_two=nrow(shop7)
p2=(more_than_or_equal_to_two/total_customers)*100
p2
```

```{r warning=FALSE}
# Frequency of average monthly visits greator or equal to 1
shop7<-filter(shop6,Visits_per_month1>=3)
more_than_or_equal_to_three=nrow(shop7)
p3=(more_than_or_equal_to_three/total_customers)*100
p3
```

```{r warning=FALSE}
# Frequency of average monthly visits greator or equal to 1
shop7<-filter(shop6,Visits_per_month1>=4)
more_than_or_equal_to_four=nrow(shop7)
p4=(more_than_or_equal_to_four/total_customers)*100
p4
```

```{r warning=FALSE}
# Frequency of average monthly visits greator or equal to 1
shop7<-filter(shop6,Visits_per_month1>=5)
more_than_or_equal_to_five=nrow(shop7)
p5=(more_than_or_equal_to_five/total_customers)*100
p5
```

Getting the final dataset which has percentage of customers according to their average monthly visits frequency 

```{r warning=FALSE}
x=c(1,2,3,4,5)
y=c(p1,p2,p3,p4,p5)
z=z=data.frame(x,y)

#Plotting the Customer proportion vs Visits frequency
ggplot(z,aes(x,y)) + 
  geom_point() +
  geom_line()+
  xlab("Visits_frequency_per_month") + 
  ylab('Proportion of customers') + ggtitle('Customer proportion vs Visits frequency')
```

*Interpretation* <br>

We see that the percentage of customers who visits the coffee shop once a month or more 
on average is 11.4%. This is our best case scenario of repeat customers. If we increase the
threshold of average monthly visits frequency to 2,3,4 etc. the percentage of customers meeting
this threshold falls with a steep slope. There are only 0.04% of customers who visit the
coffee shop equal to or more than 5 times.

*Conclusion* <br>

By looking at our customers visits frequency we’ve gained some insights into how many of 
our customers are repeat customers which can be further identified using the right frequency threshold by business. Suppose, if we take once a month as threshold to define the customer 
loyalty then we can probe these customers to increase the frequency of their visits or
increase transaction size by directing our marketing efforts toward this customer base instead 
of the whole customer base.

